---
title: "Case Study: Working With Afrobarometer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case Study: Working With Afrobarometer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

if ( file.exists("not_included", "daniel_env.R")) {
  source(file.path("not_included", "daniel_env.R"))
} else {
  source(file.path("..", "not_included", "daniel_env.R"))
}

```

```{r setup}
library(retroharmonize)
```

In this case study we harmonize data from Afrobarometer with Eurobarometer. Some elements of this vignette are not “live”, because we do not have permission to re-publish the microdata files from Afrobarometer, but you can access them [directly](http://afrobarometer.org/data/merged-data).  For reproducibility, we are storing only a small subsample from the files and the metadata. 

## Importing Afrobarometer Files

First, let’s read in the two rounds of Afrobarometer.

```{r import}
### use here your own directory
ab <- dir ( afrobarometer_dir )
afrobarometer_rounds <- file.path(afrobarometer_dir, ab)

ab_waves <- read_surveys(afrobarometer_rounds, .f='read_spss')
attr(ab_waves[[1]], "id") <- "Afrobarometer_2016"
attr(ab_waves[[2]], "id") <- "Afrobarometer_2019"
```

Create a metadata file, or a data map with `metadata_create()`.

```{r}
ab_metadata <- lapply ( X = ab_waves, FUN = metadata_create )
ab_metadata <- do.call(rbind, ab_metadata)
```

## Working with the metadata

```{r, selection, message=FALSE}
library(dplyr)

to_harmonize <- ab_metadata %>%
  filter ( var_name_orig %in% 
             c("rowid", "COUNTRY", "REGION", "withinwt") |
             grepl("trust ", label_orig ) ) %>%
  mutate ( var_label = var_label_normalize(label_orig)) %>%
  mutate ( var_label = case_when ( 
    grepl("^unique identifier", var_label) ~ "unique_id", 
    TRUE ~ var_label)) %>%
  mutate ( var_name = val_label_normalize(var_label))
```
```{r}
head(to_harmonize %>%
       select ( all_of(c("id", "var_name", "var_label"))), 10)
```

The `merge_waves()` function harmonizes the variable names, the variable labels and survey identifiers and produces a list of surveys (of class `survey()`.) The parameter `var_harmonization` must be a list or a data frame that contains at least the original variable names, the new variable names and their labels.

```{r merge}
merged_ab <- merge_waves ( waves = ab_waves, 
                           var_harmonization = to_harmonize  )

## We do not need the labels for the countries and provinces 
merged_ab <- lapply ( merged_ab, 
         FUN = function(x) x  %>%
           mutate_at ( 
             vars(all_of(c("country","province_or_region"))), 
             as_character ) )
```

## Harmonize the values

The Afrobarometer version of the trust variables is a bit different from Eurobarometer, it has 4 categories, not 2. We could just to map them into two, or give them an equi-distant numerical representation. This is what we do. The `document_survey_item()` function shows the metadata of a single variable.

```{r check}

```

... and its code table:

```{r}
document_survey_item(merged_ab[[1]]$trust_president) %>%
  knitr::kable()
```


We create a harmonization function from the `harmonize_values()` prototype function. In fact, this is just a re-setting the default values of the original function.  It makes future reference in pipelines easier, or it can be used for a question block only, in this case to variables with `starts_with("trust")`.

```{r specify}
harmonize_ab_trust <- function(x) {
  label_list <- list(
    from = c("^not", "^just", "^somewhat",
             "^a", "^don", "^ref", "^miss"), 
    to = c("not_at_all", "little", "somewhat", 
           "a_lot", "do_not_know", "declined", "inap"), 
    numeric_values = c(0, 0.33333, 0.66666, 1, 99997, 99998, 99999)
  )
  
  harmonize_values(
    x, 
    harmonize_labels = label_list, 
    na_values = c("do_not_know"=99997,
                  "declined"=99998,
                  "inap"=99999)
  )
}
```

Let's apply these settings to the trust variables. The `harmonize_waves()`  function binds all variables that are present in all surveys.

```{r harmonize}
harmonized_waves <- harmonize_waves ( 
  waves = merged_ab, 
  .f = harmonize_ab_trust ) %>%
  mutate ( year = ifelse (
    grepl("r6_data_2016", unique_id), 
    yes = 2016, no = 2019)) 
```

## Analyze the results

The numeric form of these trust variables is comparable with the numeric averages of the Eurobarometer trust variables.

```{r numeric}
harmonized_waves %>%
  mutate_at ( vars(starts_with("trust")), 
              ~as_numeric(.)*within_country_weighting_factor) %>%
  select ( -all_of("within_country_weighting_factor")) %>%
  group_by ( country ) %>%
  summarize_if ( is.numeric, mean, na.rm=TRUE)
```

And the factor presentation, without weighting:

```{r factor}
require(tidyr)
harmonized_waves %>%
  select ( -all_of("within_country_weighting_factor")) %>%
  mutate_if ( is.labelled_spss_survey, as_factor) %>%
  tidyr::pivot_longer ( starts_with("trust"), 
                        names_to  = "institution", 
                        values_to = "category") %>%
  mutate ( institution = gsub("^trust_", "", institution) ) %>%
  group_by ( country, institution, category ) %>%
  summarize ( n = n()) 
```

