---
title: 'Case Study: Working With Arab Barometer Surveys'
author:
- Ahmed Shaibani
- Daniel Antal
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette: default
vignette: >
  %\VignetteIndexEntry{Case Study: Working With Arab Barometer Surveys}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, include=FALSE}
## https://github.com/tidyverse/rvest/blob/master/vignettes/selectorgadget.Rmd
requireNamespace("png", quietly = TRUE)
embed_png <- function(path, dpi = NULL) {
  meta <- attr(png::readPNG(path, native = TRUE, info = TRUE), "info")
  if (!is.null(dpi)) meta$dpi <- rep(dpi, 2)
  knitr::asis_output(paste0(
    "<img src='", path, "'",
    " width=", round(meta$dim[1] / (meta$dpi[1] / 96)),
    " height=", round(meta$dim[2] / (meta$dpi[2] / 96)),
    " />"
  ))
}
```

```{r setup-knitr, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r vignette-setup, echo=FALSE, message=FALSE}
library(retroharmonize)
library(dplyr, quietly=TRUE)
library(tidyr)
library(ggplot2)
library(knitr)
library(here)
here::here()

load(file = system.file(
  # the first steps are saved becasue of 
  # large file/object sizes.
  "arabb", "arabb_vignette.rda", 
  package = "retroharmonize"))
```

The goal of this case study is to explore the variation in trust in various state institutions among African societies, as well as changes in trust over time.

To do this, we use data from [Arab Barometer](https://https://www.arabbarometer.org/), a nonpartisan research network that provides insight into the social, political, and economic attitudes and values of ordinary citizens across the Arab world. 

`retroharmonize` is not affiliated with Arab Barometer. To fully reproduce this example, you must acquire the data files from them, which is free of charge. If you download and use the survey data, the Arab Barometer does not take any responsibility for the results of the analysis of the Arab Barometer data. All results published using the Arab Barometer data and related content are responsibility of the respective author.

Some elements of the vignette are not “live”, because we want to avoid re-publishing the original microdata files from Arab Barometer. You can access the data directly from the [arabbarometer.org website](https://www.arabbarometer.org/survey-data/data-downloads/) after agreeing to their [policies](https://www.arabbarometer.org/policies/) and filling out a short questionnaire.

The files are stored on the website in `.zip` format, which may be extracted to folders. We assume that you extracted and copied all `.sav` files into a single folder that we will call in this vignette the `arabbarometer_dir`. Define your own `arabbarometer_dir` with `file.path()` in your own system.

  
File name            | Wave
-------------------- | -------------
ABI_English.sav      | Arab Barometer Wave I
ABII_English.sav     | Arab Barometer Wave II
ABIII_English.sav    | Arab Barometer Wave III
ABIV_English.sav     | Arab Barometer Wave IV
AB-WaveV-EN.sav      | Arab Barometer Wave V


## Importing Arab Barometer Files

We start by reading in the three rounds of the Arab Barometer. The following chunk is not evaluated, if you want to re-run this tutorial, you need to define your `arabbarometer_dir` and download the Arab Barometer files there, after acknowledging Arab Barometer’s terms and conditions. Alternatively, you can load the pre-saved results of the first steps in the first, `vignette-setup` chunk.

```{r import, eval=FALSE}
### use here your own <arabbarometer_dir> directory
ab <- dir(arabbarometer_dir, pattern = "sav$")
arabbarometer_rounds <- file.path(arabbarometer_dir, ab)
arab_waves <- read_surveys(arabbarometer_rounds, .f='read_spss')
```

we appended meaningful identifiers to the file names as follows (you must follow the ordering of your files in your own `arabbarometer_dir` dir:

```{r wave-id, eval=FALSE}
# In the vignette this is pre-loaded.
attr(arab_waves[[1]],"id") <- "Arab_Barometer_5"
attr(arab_waves[[2]],"id") <- "Arab_Barometer_1"
attr(arab_waves[[3]],"id") <- "Arab_Barometer_2"
attr(arab_waves[[4]],"id") <- "Arab_Barometer_3"
attr(arab_waves[[5]],"id") <- "Arab_Barometer_4"
```

### Importing Arabbarometer Datasets

We used `document_waves()`to see the primary descriptive information of our survey wave (one harmonized survey set) files. Visually check that the file names correspond with the meaningful IDs that you gave to each survey wave, i.e., `Arab_Barometer_1` corresponds with `ABI_English.sav`.

```{r document-arab-waves, eval=FALSE}
# In the vignette this is pre-loaded.
documented_arab_waves <- document_waves(arab_waves)
```

```{r print_documented_arab_waves}
print(documented_arab_waves)
```
We obtained 5 nested lists in the `arab_waves` list, each of them containing data of one survey wave Arab Barometer. To create longitudinal tables, i.e., join several survey waves into a single table, it is important to review the methodology of each round of Arabbarometer’s surveys, its structure, modifications, and changes in topics of interest.

### Merged Metadata Creation 

In this step, we built a comprehensive metadata table that maps information about variable names and labels of the case’s surveys data. Each row in the table above refers to a different variable in the survey data file. This is accomplished using the `metadata_create()` function, which collects metadata from survey data files, normalizes variable labels, and defines substantive response ranges and missing value codes.


```{r create-arabb-metadata, eval=FALSE}
# In the vignette this is pre-loaded.
arabb_metadata <- lapply ( X = arab_waves, FUN = metadata_create)
arabb_metadata <- do.call(rbind, arabb_metadata)
```

Let's randomly review some metadata:

```{r random-review}
set.seed(2021)
arabb_metadata %>% 
  select (-all_of(c("filename", "class_orig"))) %>%
  sample_n(10)
```


We select from the metadata file `rowid` being the unique case identifier, and variables of interest:  `date` with the interview date,`COUNTRY` containing information about the country where the interview was conducted, and with `wt` being the weighting factor, and some "real" survey questions.

In the same context, you should also carefully insert key words and expressions that match with your survey questionnaire item of interest; in our case, we chose *Election, parliamentary*, and *Election* to fork an exact question about parliamentary election, which is `How would you evaluate the last parliamentary election?`

The variable names (`var_name`) and labels (`var_label`) then must be normalized and harmonized.

### Harmonizing the variable names

```{r arabb-metadata}
to_harmonize <- arabb_metadata %>%
  filter( .data$var_name_orig %in% c("rowid", "country","date", "wt")|
           grepl("how would you evaluate the last parliamentary", .data$label_orig)) %>%
  mutate(var_label = var_label_normalize(.data$label_orig)) %>%
  mutate(var_label = case_when(
    .data$var_name_orig == "country" ~ "Country",
    .data$var_name_orig == "rowid"   ~ "Unique ID AB English", # in pdf Unique ID AB English
    .data$var_name_orig == "date"    ~ "Date_of_interview",
    .data$var_name_orig == "wt"      ~ "Weight",
    TRUE ~ " Evaluation in the last parliamentary elections")) %>%
  mutate ( var_name = var_label_normalize(.data$var_label) )

set.seed(2021) # Let's see the same random example:
sample_n(to_harmonize%>% 
       select ( all_of(c("id", "var_name", "var_label"))), 10)
```

The function `merge_waves()` harmonizes the variable names, the variable labels and survey identifiers and returns a list of surveys (of `class survey()`.) The parameter _var_harmonization_ must be a list or a data frame that contains at least the original file name (`filename`), original variable names (`var_name_orig`), the new variable names (`var_name`) and their labels (var_label), so that the function correctly makes the harmonization steps.

```{r merge-arabb-waves, eval=FALSE}
# In the vignette this is pre-loaded.
merged_ab <- merge_waves(waves = arab_waves, 
                         var_harmonization = to_harmonize)
merged_ab <- lapply (merged_ab, 
                     FUN = function(x) x  %>%
                       mutate( country = as_character(country)))
```


```{r save-first-steps, include=FALSE}
# This saves the first steps because the object sizes are too large to be included with the package on CRAN.
save (merged_ab, arabb_metadata, documented_arab_waves, file = here("inst", "arabb", "arabb_vignette.rda"))
```


```{r clean-country-names}
# cleaning country names using the second round
merged_ab[[2]]$country <- gsub("\\d{1,}\\.\\s?","",merged_ab[[2]]$country)
merged_ab[[2]]$country <- trimws(merged_ab[[2]]$country, which="both")
```

```{r document-merged-ab}
documented_merged_ab <- document_waves(merged_ab)
print(documented_merged_ab)
```
 
Let's take a look at the variables: 

```{r var-names}
lapply (merged_ab, names)
```
We see that they are not set in a standard format (variable names should not have spaces for programmatic use). We apply snakecase variable naming:

```{r snakecase, message=FALSE}
require(snakecase)
# replace space with underscore form variable names
merged_ab <- lapply(merged_ab,
                    FUN = function(df) df %>%
                      rename_all(snakecase::to_snake_case))
```

Now the variable names are ready to use in the programmatic way:

```{r var-names-2}
lapply(merged_ab, names)
```
As shown from the table above, we obtained a reduced form of a comprehensive metadata in nested data set format. 

It is useful to check the variable names in each round, and filter out nonessential variables and if the full round data does not have the question of our interest, the omitting the round is possible.

In our case, we decided to:
- omit completely both waves 1 and 5, because none of them has our key variable `Parliamentary Election Evaluation`;
- omit `date of interview` as it is missing in both waves (rounds) 2 and 4.


```{r remove-waves}
merged_ab[[1]] <- NULL # Remove AB-WaveV-EN_1
merged_ab[[2]] <- NULL # Remove ABI_English
```

```{r remove-date, eval=TRUE}
merged_arabb <- lapply(merged_ab, function(x){
  if ("date_of_interview" %in% names(x)){
    subset(x, select = -c(date_of_interview))
  } else{
    subset(x)
  }
})
```

```{r print-remaining-waves}
document_waves(merged_arabb)
```

### Variable Label Harmonization 

The question of interest "how would you evaluate the last parliamentary election?" is described as a closed form survey questions where answers are predetermined and the respondents chose one of the answer options. This question has four-point ordinal scales, therefore, data is best to be analyzed with ordinal models which do not assume that the points are equidistant, i.e. as factors.

In order to get a quick idea of how the data look like, we assigned numbers 1-4 as follows:

1. Answer with 1 corresponding to the `free_and_fair` (highest democratic election process).
2. Answer with 2 corresponding to `some_minor_problems`.
3. Answer with 3 corresponding to `some_major_problems`.
4. Answer with 4 corresponding to `not_free`(lowest democratic election process).
5. Answer with other values like 0 and higher than 4 is classified as either missing answers, do not know or refuse/declined to answers.


### Harmonizing the Missing Values

Arabbarometer's SPSS files use various missing value categories, wo we have to carefully harmonize them, too. The entire value range is the following:

```{r review-value-labels, eval=TRUE}
collect_val_labels(
  to_harmonize %>% 
     filter(grepl("evaluation in the last parliamentary elections", 
                  .data$var_name))
  )
```

>It is crucial to draw attention to special cases when some `NA` values are not labeled profoundly in data tables, thus, we decided to replace `NA` in advance with the value of the "missing/or not applicable" before standardizing response's values. Notice that we assigned original `NA` or missing labels with numerical values as in the table below:

```{r na-coding}

missing_value_labels <- c("do_not_know","declined","declined","missing")
missing_value_codes <- c("99997","99998","99998","99999")

na_harmonized_df <- data.frame(
  original_label = missing_value_labels,
  harmonized_values = missing_value_codes,
  stringsAsFactors = FALSE)

na_harmonized_df
```
For safety, we need to handle the `NA` values before harmonization (this is important), as function harmonize_arabb_trust could not distinguish NA if not labeled. so it is better to replace N/A immediately with the value of the "missing/or not applicable" before standardizing response' values. we could add label to N/A after adding the values but her not necessary. 

```{r harmonize-na-values, eval=FALSE}
## This is not code
merged_arabb <- lapply(merged_arabb, function(df) {
  df$evaluation_in_the_last_parliamentary_elections[
    is.na(df$evaluation_in_the_last_parliamentary_elections)==T] <- 99999
  return(df)
} )
# merged_arabb[[3]]$evaluation_in_the_last_parliamentary_elections[is.na(merged_arabb[[3]]$evaluation_in_the_last_parliamentary_elections)==T]<-99999 
# merged_arabb[[1]]$evaluation_in_the_last_parliamentary_elections[is.na(merged_arabb[[1]]$evaluation_in_the_last_parliamentary_elections)==T]<-99999 
# merged_arabb[[2]]$evaluation_in_the_last_parliamentary_elections[is.na(merged_arabb[[2]]$evaluation_in_the_last_parliamentary_elections)==T]<-99999 
```

The harmonization function from the `harmonize_values()` is a prototype function. In fact, this is just a re-setting the default values of the original function. It makes future reference in pipelines easier, or it can be used for a question block only, in this case to variables the are correlated to  our question about ("Evaluation of the Parliamentary Election").


```{r harmonize-arabb-trust}
harmonize_arabb_trust <- function(x){
  label_list <- list(
    from = c("(\\d\\.\\s)?(\\w+\\s\\w+\\s)?([c|C]ompletely free and fair)",
             "(.+)(but)?\\s?(with)\\s(some)?\\s{0,}(minor\\s\\w+)",
             "(.+)(but)?\\s?(with)\\s(some)?\\s{0,}(major\\s\\w+)",
             "(.+)?([n|N]ot\\sfree\\s\\w+\\s\\w+)",
             "((\\d.\\s{0,})?\\si\\s)?([d|D]on.t\\sknow)(\\s\\(Do\\snot\\sread\\))?", 
             "[R|r]efuse", 
             "(\\d.\\s)?[d|D]ecline[d]?(\\s\\w+\\s\\w+)(\\s.Do not read.)?",
             "(\\d.\\s)?[m|M]issing"),
    to = c("free_and_fair", 
           "some_minor_problems",
           "some_major_problems",
           "not_free",
           "do_not_know","declined","declined","missing"),
    numeric_values = c(3,2,1,0,99997,99998,99998,99999))
  harmonize_values(x, harmonize_labels = label_list, 
                   na_values = c("do_not_know"= 99997,
                                 "declined"=99998,
                                 "missing"=99999
                   ))
}
```

```{r harmonize-arabb-waves, eval=TRUE}
harmonized_arabb_waves <- harmonize_waves( 
  waves = merged_arabb, 
  .f = harmonize_arabb_trust )
```

We tested the transformed variables after applying the harmonization function, and we obtained the following counting summary of each unique answer of our question. 

```{r, eval=TRUE}
harmonized_arabb_waves %>%
  mutate ( weight = ifelse(is.na(.data$weight), 1, .data$weight), 
           evaluation_in_the_last_parliamentary_elections = as_factor(evaluation_in_the_last_parliamentary_elections), 
           country = as_factor(country) ) %>%
  select ( -.data$unique_id_ab_english ) %>%
  summary()
```
 
### Analyzing the harmonized survey's data of parliamentary election in MENA

From the attributes of the harmonized_ab_waves, we observed unnecessary information that could be omitted such as row names. 

```{r, eval=TRUE}
h_ab_structure <- attributes(harmonized_arabb_waves)
h_ab_structure$row.names <- NULL 
```

The harmonized data can be exported and analyzed in another statistical program. 

The labeled survey data was stored in `labelled_spss_survey()` vectors, which is a complex class that retains metadata for reproducibility, but most statistical packages work with numeric or categorical (factor) data. The data can be converted to any of these, but you should use the methods of `labelled_spss_survey()` for this purpose, i.e., `as_numeric()` or `as_factor()`. (See more why you should not fall back on the more generic as.factor() or as.numeric() methods in The [labelled_spss_survey class vignette](https://retroharmonize.dataobservatory.eu/articles/labelled_spss_survey.html).)

Due to the categorical format of the question of MENA parliamentary election evaluation in the Arab Barometer survey rounds, we decided to present the election variable in its *categorical format*, and disregarded the post-stratification weights.

```{r simply-wave-names, eval=TRUE}
harmonized_arabb_waves <- harmonized_arabb_waves %>%
  mutate(wave = stringr::str_extract(unique_id_ab_english, "(\\b[A-Z0-9]+)"))
```

```{r valuation-summary, eval=TRUE}
categorical_summary <- harmonized_arabb_waves %>%
  select ( -all_of(c("weight", "unique_id_ab_english")) ) %>%
  mutate ( 
    evaluation_in_the_last_parliamentary_elections = 
      as_factor(evaluation_in_the_last_parliamentary_elections)) %>%
  pivot_longer ( starts_with("evaluation"), 
                 names_to  = "indicator", 
                 values_to = "valuation") %>%
  group_by ( .data$country, .data$wave, .data$valuation ) %>%
  summarize (n = n(),
             .groups = 'keep') 
```

```{r print-summary, eval=TRUE}
categorical_summary[1:10,]
```

The output as shown in the table below indicates the number of votes that were given to each category of the surveys' parliamentary election question in country and round levels. For instance, the output of the arabbarometric survey round 2 in Algeria shows a high level of dissatisfaction of the last election process with 273 respondents expressed that it was not free, and 155 voters said it had significant problems, while 287 mentioned to some minor problems in the election process, and only 76 opinions indicated that the election process was fair. Furthermore, we notice the high number of 'do not know' answers in this round which indicates to some hesitation in respondents' willingness to express their opinions.This was an example of one country.  

### Regional Analysis of MENA Parliamentary Election Freedom Evaluation 

```{r regional_parl_elections-2, eval=TRUE, message=FALSE}
regional_parl_elections <- harmonized_arabb_waves %>%
  select(
    -all_of(c("weight", "unique_id_ab_english"))
    )%>%
  mutate ( 
    evaluation_in_the_last_parliamentary_elections = 
      as_factor(evaluation_in_the_last_parliamentary_elections)) %>%
  pivot_longer ( starts_with("evaluation"), 
                 names_to  = "indicator", 
                 values_to = "valuation") %>%
      group_by(.data$valuation) %>%
  summarize(n=n()) 
```

```{r, eval=TRUE, message=FALSE}
regional_parl_elections[1:5,]

regional_not_free <- regional_parl_elections%>%
  filter(.data$valuation == 'not_free')
regional_maj_prob <- regional_parl_elections %>%
  filter(.data$valuation == 'some_major_problems')
regional_min_prob <- regional_parl_elections %>%
  filter(.data$valuation == 'some_minor_problems')
regional_free <- regional_parl_elections %>%
  filter(.data$valuation == 'Free_and_fair')
regional_dont_know <- regional_parl_elections %>%
  filter(.data$valuation == 'do_not_know')

```

Starting with mentioning that round ABII of the survey has sample size of 12782 voters compared to 14809 of round ABIII and 9000 of round ABIV. We summarized the level of democracy of the parliamentary election process or voters' satisfaction in MENA region to three tables (see below).

The low democracy level/low satisfaction about the fairness of the election process is presented in the table `problama_Elec_process `. The total votes that described the process as *not-free* or *with major problems* decreased significantly from 44% in 2007-round ABII to less than 30% in 2010-Round ABIV using the total_unfair variable. 


```{r regional_parl_elections, eval=TRUE}
## needs review
problem_elec_process <- bind_rows(
  regional_not_free,
  regional_maj_prob
  )

problem_elec_process <- problem_elec_process %>% 
  group_by(.data$valuation) %>% 
  mutate(total_unfair = sum(.data$n)) %>% 
  arrange( .data$valuation, .data$total_unfair)

print(problem_elec_process)
```


On the other side, the table`Good_Elec_process` indicates to a temporarily improvement in absolute number of votes in the round ABIII  compared to the round ABII before returning to same level in round ABII, but in percentage scale, the people who expressed that the election was either *free and fair* or *with minor problems* increased from 32% in 2010 - round ABII to 47% in 2017- round ABIV using total_fair variable. 


```{r, eval=TRUE}

Good_Elec_process <- bind_rows(
  regional_free,
  regional_min_prob
  )

#Good_Elec_process <- Good_Elec_process %>% 
#  group_by(.data.$valuation) %>% 
#  mutate(total_fair = sum(n)) %>% 
#  arrange(.data$valuation,.data$total_fair)
print(Good_Elec_process)
```

The rest of the people who responded to this question with "I do not know" increased from 11.3% in 2007 to 15.6% in 2010. 


```{r, eval=TRUE}
reluctant_respondents <- regional_dont_know
print(reluctant_respondents)
```

Eventually, users can go deeper in analyzing the harmonized data of our question about the parliamentary election process in MENA, and they can test the library on other questions to reach and explore other interesting results. The main purpose of this blog was to show the power of **retroharmonize** in handling large and complex survey data in a way that help users to obtain the required output effectively. 

The `merge_waves()` function harmonizes the variable names, the variable labels and survey identifiers and returns a list of surveys (of class `survey()`.) The parameter `var_harmonization` must be a list or a data frame that contains at least the original file name (`filename`), original variable names (`var_name_orig`), the new variable names (`var_name`) and their labels (`var_label`), so that the program knows which variables to take from what files and how to call and label them after transformation.


```{r merge}
merged_ab <- merge_waves ( waves = arab_waves, 
                           var_harmonization = to_harmonize  )

# country will be a character variable, and doesn't need a label
merged_ab <- lapply ( merged_ab, 
         FUN = function(x) x  %>%
           mutate( country = as_character(country)))
```

Review the most important metadata with `document_waves()`:

```{r}
documenteded_merged_ab <- document_waves(merged_ab)
```
```{r}
print(documenteded_merged_ab)
```

The files we want to harmonize must contain the same variables. 
If we want to weight the answers, we should either drop Arab-Barometer_1, or for simplicity, add weight = 1 to this survey.

```{r, eval=FALSE}
## not evaluated, this is how to drop the Arab-Barometer_1
merged_ab[[2]] <- NULL
```


```{r add-weight}
merged_ab[[2]]$weight <- 1
```

Similarly, we can either drop the date of the interview or fill it up with, for example, the middle day of the surveying period.

```{r drop-date}
merged_arabb <- lapply ( merged_ab, function(x){
  if ( "date_of_interview" %in% names(x) ) {
    subset ( x, select = -c(date_of_interview))
  } else {
    ## subset works with survey class, if omitted, will return 
    ## data frame without metadata
    subset ( x )
  }
})
```

```{r}
documenteded_merged_arabb <- document_waves(merged_arabb)
```
```{r}
print(documenteded_merged_arabb)
```

Now we have the same four variables in each wave: `r paste ( names(merged_arabb[[1]]), collapse = ", ")`.

## Harmonization

To review the harmonization on a single survey use `pull_survey()`. Here we select Arab Barometer Round 3.

```{r check}
R3 <- pull_survey ( merged_ab, id = "Arab_Barometer_3")
```

```{r pulled-attributes}
attributes(R3$evaluation_economic_situation[1:20])
```

The `document_survey_item()` function shows the metadata of a single variable. 

```{r document-item}
document_survey_item(R3$evaluation_economic_situation)
```

Arab Barometer's SPSS files do not mark the missing values, so we have to be careful. 

The set of valid category labels and missing value labels are as follows:
  
```{r}
collect_val_labels (to_harmonize %>%
                      filter ( grepl( "evaluation_economic_situation", var_name) ))
```

Apparently, there are a number of missing value labels here, and it is not even always clear if they refer to the same type of “missingness” with the same word.

In the examples with [Afrobarometer](https://retroharmonize.dataobservatory.eu/articles/afrobarometer.html) and [Eurobarometer](https://retroharmonize.dataobservatory.eu/articles/eurobarometer.html), we saw different problems with missing variables.

* In Eurobarometer, the type of missingness is relatively clear, and in the SPSS files, the missing labels are usually, but not always identified.
* In Afrobarometer, the missing classes are clear, but the SPSS missing labels are not used. Instead the SPSS missing range (`na_range`) is used, but sometimes inconsistently with the missing labels (`na_labels`).

In some statistical analysis, these differences do not matter much (for example, when creating averages from numeric values or numeric representations), but when you work with categories or you examine the survey's quality, these differences can be important.

We create a harmonization function from the `harmonize_values()` prototype function. In fact, this is just a re-setting the default values of the original function.  It makes future reference in pipelines easier, or it can be used for a question block only, in this case to variables with `starts_with("trust")`.

```{r specify}
harmonize_arabb_trust <- function(x) {
  label_list <- list(
    from = c("very\\sgood", "(?<!very\\s)good",
             "(?<!very\\s)bad", "very\\sbad", 
             "t\\sknow", "refuse", "decline", "missing"), 
    to = c("very_good", "good", "bad", 
           "very_bad", "do_not_know", 
           "declined", "declined", "missing"), 
    numeric_values = c(3,2,1,0, 99997, 99998, 99998,99999)
  )
  
  harmonize_values(
    x, 
    harmonize_labels = label_list, 
    na_values = c("do_not_know"=99997,
                  "declined"=99998,
                  "missing"=99999), 
    perl = TRUE
  )
}
```

Let's see if the output makes sense: 

```{r, eval=FALSE}
# review this
# We only print a random sample from the results.
set.seed(2020)
harmonize_arabb_trust(
  sample(R3$evaluation_economic_situation, 12)
)
```

We apply this function to the trust variables. The `harmonize_waves()` function binds all variables that are present in all surveys.

```{r harmonize, eval=FALSE}
harmonized_arabb_waves <- harmonize_waves ( 
  waves = merged_arabb, 
  .f = harmonize_arabb_trust ) 
```

Let's look at the attributes of `harmonized_arabb_waves`. 

```{r, eval=FALSE}
h_arabb_structure <- attributes(harmonized_arabb_waves)
```
```{r, eval=FALSE}
h_arabb_structure$row.names <- NULL # We have over 70K row names
h_arabb_structure
```

To keep our example manageable, we subset the datasets to include only five countries.

```{r, eval=FALSE}
harmonized_arabb_waves <- harmonized_arabb_waves %>% 
  mutate ( id = strsplit(unique_id_ab_english, "[_]")[[1]][1])
```


## Analyzing the harmonized data

The harmonized data can be exported and analyzed in another statistical program.  The labelled survey data is stored in `labelled_spss_survey()` vectors, which is a complex class that retains metadata for reproducibility. Most statistical R packages do not know it. The data should be presented either as numeric data with `as_numeric()` or as categorical with `as_factor()`.  (See more why you should not fall back on the more generic `as.factor()` or `as.numeric()` methods in [The labelled_spss_survey class vignette.](https://retroharmonize.dataobservatory.eu/articles/labelled_spss_survey.html))

Please note that the numeric form of these trust variables is not directly comparable with the numeric averages of the Eurobarometer trust variables, because the middle of the range is at `r mean(0:3)` and not `r mean(0:1)`.

```{r numeric}
numeric_summary <- harmonized_arabb_waves %>%
  mutate_at ( vars(starts_with("evaluation")), 
              ~as_numeric(.)*weight) %>%
  select ( -all_of("weight") ) %>%
  group_by ( .data$country ) %>%
  summarize_if ( is.numeric, mean, na.rm=TRUE )

numeric_summary
```

And the factor representation, without weighting:

```{r factor, eval=FALSE}
library(tidyr)  ## tidyr::pivot_longer()
categorical_summary <- harmonized_arabb_waves %>%
  select ( -all_of(c("weight", "unique_id_ab_english")) ) %>%
  mutate ( 
    evaluation_economic_situation = 
      as_factor(evaluation_economic_situation)) %>%
  pivot_longer ( starts_with("evaluation"), 
                        names_to  = "indicator", 
                        values_to = "valuation") %>%
  group_by ( .data$country, .data$valuation ) %>%
  summarize ( n = n() ) 

categorical_summary
```

### Saved Variables

```{r, eval=FALSE}
haven::write_sav(data = harmonized_arabb_waves, "harmonized_arabb_waves.sav")
write.csv(harmonized_arabb_waves, "harmonized_arabb_waves.csv")
write.csv(categorical_summary, "categorical_summary.csv")
write.csv(numeric_summary, "numeric_summary.csv")

# The metadata file contains list objects, which cannot be represented
# in a flat csv file format.
saveRDS(arabb_metadata, "arabb_metadata.rds")

## The lists of value labels are dropped from the csv output.
write.csv(arabb_metadata [, -which (sapply ( arabb_metadata, class) == "list")], "arabb_metadata_simplified.csv")
```


