---
title: "Working With Arabbarometer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working With Arabbarometer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`retroharmonize` is not affiliated with Arabbarometer.  For this example, you need to download and store in a directory the following files from their website [here](addurl) after carefully reading their terms of use: `ABI_English.sav`      `ABII_English.sav`, `ABIII_English.sav`, `ABIV_English.sav`, `ABV_Release_Data.sav`.  We will refer to the library that contains these files as `arabbarometer_dir`.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



```{r setup}
library(retroharmonize)
library(tidyverse)
```

```{r readfiles, eval=FALSE}
arabbarometer_dir <- file.path("not_included", "arabb", "input")
ab <- dir (arabbarometer_dir, pattern = "sav$")
arabbarometer_rounds <- file.path(arabbarometer_dir, ab)
arab_waves <- read_surveys(arabbarometer_rounds,.f='read_spss')
attr(arab_waves[[1]],"id") <- "Arab_Barometer_1"
attr(arab_waves[[2]],"id") <- "Arab_Barometer_2"
attr(arab_waves[[3]],"id") <- "Arab_Barometer_3"
attr(arab_waves[[4]],"id") <- "Arab_Barometer_4"
attr(arab_waves[[5]],"id") <- "Arab_Barometer_5"

documented_arab_waves <- document_waves(arab_waves)
save(documented_arab_waves, file = file.path("data-raw", "arabb", "Arab.Rda"))

```

