---
title: "Creating a Harmonized Cultural Access & Participation Dataset for Music"
author: "Daniel Antal"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating a Harmonized Cultural Access & Participation Dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitrsetup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, eval=TRUE,
  comment = "#>"
)
retroharmonize::here()
```

```{r setup, message=FALSE, warning=FALSE}
library(retroharmonize)
library(dplyr, quietly = TRUE)
cap_files <- c('ZA4529_v3-0-1.sav', 'ZA5688_v6-0-0.sav')
```

We are going to create a harmonized dataset from two Cultural Access & Participation datasets, which each contain harmonized surveys from a 2007 and 2013.  We will further harmonize them into a single, longitudional file.

* European Commission (2012). Eurobarometer 67.1 (Feb-Mar 2007). GESIS Data Archive, Cologne. ZA4529 Data file Version 3.0.1, [https://doi.org/10.4232/1.10983](https://doi.org/10.4232/1.12959).

* European Commission, Brussels (2016). Eurobarometer 79.2 (2013). GESIS Data Archive, Cologne. ZA5688 Data file Version 6.0.0, [https://doi.org/10.4232/1.12577](https://doi.org/10.4232/1.12577).

To replicate this example, you must acquire these files from [GESIS](https://www.gesis.org/home) after accepting their terms of use.  The `retroharmonize` project is not affiliated with GESIS.

This vignette article shows you how to replicate our dataset. The results can be found and referenced as:
* Daniel Antal. (2022). Harmonized Cultural Access & Participation Dataset for Music (Version 20220002) [Data set]. Zenodo. [https://doi.org/10.5281/zenodo.5917742](10.5281/zenodo.6606867)

* Daniel Antal. (2022). Creating a Harmonized Cultural Access & Participation Dataset for Music (Version 20220129) [Data set]. Zenodo. [10.5281/zenodo.5917714](https://doi.org/10.5281/zenodo.5917714)

```{r, eval=FALSE}
## This is a dummy path. You should use the path where you saved these files.
gesis_dir <- file.path("C:", "your_data", "gesis_files")
```

```{r loadgesisdir, inclue=FALSE}
source(retroharmonize::here("not_included", "daniel_env.R"))
```

```{r createpath, eval=TRUE}
## Create full paths to the three selected files
cap_files <- retroharmonize::here(gesis_dir, c('ZA4529_v3-0-1.sav', 'ZA5688_v6-0-0.sav'))
```

In this example, we will create the following harmonized variables in a single harmonized data table:

- Combinations from two important demography variables: the age of the respondent, and the age of school leaving of the respondent (which is a Eurobarometer-specific education level proxy.)
- Two joint weight variables from the sub-sample post-stratification weights and the subsample-independent projection weights.

The Eurobarometer master questionnaire is made in English and French, however, these working documents are not asked from respondents.  For example, in Ireland, a localized English version is used, and in France, a localized French questionnaire is presented to the respondents. Simple questionnaire items are not changed---the ones we will use in this case are the same on the basic and Irish questionnaire. In non-English and non-French speaking regions local translations as are asked.

- [ZA5688_bq.pdf Basic Questionnaire (English & French) (Questionnaire) 974.06 KB](https://access.gesis.org/dbk/52187)---the master English/French questionnaire
- [ZA5688_q_ie.pdf Ireland (Questionnaire) 94.25 KB](https://access.gesis.org/dbk/52207)---the Irish English questionnaire
- [ZA5688_q_hu.pdf Hungary (Questionnaire) 157.52 KB](https://access.gesis.org/dbk/52206)---the Hungarian questionnaire.

The coding of the survey items are the same in all subsamples.  The SPSS files follows the labelling of the Basic questionnaire, in some cases, with typos, and missing up English and French spelling. The not-normalized labelling makes working with several files challenging.

## Import and Inventory

It is likely that the surveys will fit into your computer's memory, so you can omit the first step. If you work with many files, and you want to keep working sequentially with survey files, it is a good idea to convert them to R objects. 

```{r importsurveys}
cap_surveys <- read_surveys(cap_files)
```

```{r inventory}
survey_doc <- document_surveys(survey_list = cap_surveys)
survey_doc
```

The files are rather large. We have `sum(document_cap_files$nrow)` of `sum(document_cap_files$ncol)` variables.  We will only harmonize a few of them.

## Concepts to harmonize

```{r metadatasurveys}
cap_metadata <- metadata_create(survey_list = cap_surveys)
```
```{r metadatasurveysprint}
set.seed(6)
cap_metadata %>%
  sample_n(6)
```


## Harmonization of Variables

We will harmonize weight, socio-demography and cultural access variables. 

### Weights

From the metadata description, we select the post-stratification weight variables and projection weights. The `wex(tra)` variables are projection weights, the others' are weights. Unfortunately, they are not uniformly named in different 'waves' of the Eurobarometer survey SPSS files.

```{r weightvars}
weight_variables <- cap_metadata %>% 
  filter ( .data$var_name_orig %in% c("isocntry", "wex", "wextra", "v47", "v7", "w1") | 
             .data$var_label_orig %in% c("w_1_weight_result_from_target", 
                                         "w_3_weight_special_germany",
                                         "weight_result_from_traget_united_germany", 
                                         "w_4_weight_special_united_kingdom", 
                                         "weight_result_from_traget_united_kingdom"))
```

A *schema crosswalk* is a table that shows equivalent elements (or "fields") in more than one structured data source. With `crosswalk_table_create()` we first create an empty schema crosswalk, then we fill up the empty schema with values.  Researchers who feel more comfortable working in a spreadsheet application can create a similar crosswalk table in Excel, Numbers, or OpenOffice, and import the data from a `csv` or any tabular file.

```{r weightcrosswalk}
## See https://github.com/rOpenGov/retroharmonize/issues/21
weigthing_crosswalk_table <- crosswalk_table_create(
  ## Create an empty schema crosswalk for the weight variables 
  weight_variables
  ) %>%
  mutate ( 
    # Define the new, harmonized variable names
    var_name_target = case_when (
      # grepl("weight_result_from_target", .data$val_label_target)  ~ "w1",  [this is the issue]
      .data$var_name_orig  %in% c("wex", "wextra", "v47")        ~ 'wex',
      .data$var_name_orig  %in% c("w1", "v8")                    ~ "w1",
      .data$var_name_orig %in% c("w3a", "v12")                   ~ "w_de",
      .data$var_name_orig %in% c("w4a", "v10")                   ~ "w_uk",
      .data$var_name_orig == "rowid"  ~ 'rowid',  # do not forget to keep the unique row IDs
      TRUE ~ "geo"), 
    # Define the target R class for working with these variables.
    class_target = ifelse(.data$var_name_target %in% c("geo", "v47"), "factor", "numeric")
    ) %>%
  select ( 
    -all_of(c("val_numeric_orig", "val_numeric_target", "val_label_orig", "val_label_target"))
    ) 
```

The crosswalk table contains the original (source) variable names that must be converted to the target variable names, i.e. `v7` and `isocntry` to `geo` and `v47`, `wextra` to `wex`. The weights should remain numeric variables, and the `geo` variable should be a categorical factor variable.

```{r harmonizationweigths}
weigthing_crosswalk_table
```

For the variable `geo` the appropriate post-stratification weight is the `w1` variable. 

The Eurobarometer surveys contain separate samples for the former West Germany (`DE-W`), East Germany (`DE-E`), Northern Ireland (`GB-NIR`), and Great Britain, i.e. England, Scotland and Wales (`GB-GBN`). For the two German subsamples, it is `w_de`, and for the two United Kingdom subsamples it is the `w_uk`.  We create two new variables, the `country_code` and `w`. When the two UK and the two German samples are joined, the appropriate post-stratification weight is `w`.

```{r createweights}
weight_vars <- crosswalk(survey_list = cap_surveys, 
                         crosswalk_table =  weigthing_crosswalk_table, 
                         na_values = NULL)
```

```{r printweights}
set.seed(2022)
weight_vars %>% sample_n(6)
```

You must use the `w1` weights when you work with the `geo` variable---when you work with the separate subsamples for `GB-GBN`, `GB-NIR`, `DE-E` and `DE-W`. In a unified dataset, however, we must create an approproripate weight common weight. We create the appropriate weight variable `w` to be used with ISO country codes, i.e. `GB` and `DE`. 

```{r weightwars}
weight_vars <- weight_vars %>%
  mutate (country_code = substr(geo, 1,2),
          w = case_when ( 
            country_code == "DE" ~  w_de,  # Unified Germany post-stratification weight
            country_code == "GB" ~  w_uk,  # UK = Great Britain + Northern Ireland 
            TRUE ~ w1 )) %>%
  mutate (year_survey = case_when(
    .data$id == "ZA4529_v3-0-1" ~ '2007',
    .data$id == "ZA5688_v6-0-0" ~ '2013'
  )) %>%
  mutate (year_survey = as.factor(.data$year_survey))
```

*Beware that the country codes for the United Kingdom and for Greece follow the ISO standard, i.e. GB and GR, not the Eurostat country codes UK and EL*.

```{r printweigthvars}
weight_vars <- weight_vars %>% 
  select ( all_of(c("rowid", "country_code", "geo", "w", "w1", "wex", "id")) ) 
set.seed(2022)
weight_vars %>% sample_n(6)
```

### Demography

We will use two important demography variables: the age of the respondent, and the age of school leaving of the respondent.  This latter, `age_education` variable is an important, *ex ante* harmonized variable in Eurobarometer. Because each European country has different education systems, furthermore, because education systems are not the same across different demographic groups (for example, people schooled before the World War II often went to very different schools), a more precise education level would require a very complicated mapping with education-specific knowledge.

The `age_education` variable has three specially coded values: 
1. the value for `Declined` answers, which should be coded in a numeric representation to the special `NA` value or R;

2. An integer outside of the scope of the the valid range of responses that means the person is `Still studying`.  We will recode the still studying answers to 0, and later we will further recode them to the current age of the person.  We will also mark this special group as students---in their case, the `age_education` has a different meaning from people who already finished their education.  A 17 year-old respondent in Europe is likely to study further; a middle aged person who replied with 17 has a low education level according to the current norms. In this case, a student’s response of 17 is not the same as the middle aged person’s;

3. And at last, the special value `No formal education` means that the person did not finish the primary school.  Different countries define differently the minimum mandatory age when a person can leave the school system. We code these answers to 14, which is lower than the lowest age in our sample (15). This is also a special group, so we will mark them with a dummy variable, too.

```{r demographycrosstable}
demography_crosstable <- cap_metadata %>%
  filter ( .data$var_name_orig == "rowid" | 
           .data$var_label_orig %in% c("age_exact", "age_education")) %>%
  crosswalk_table_create() %>%
  mutate ( var_name_target = case_when(
    # Do not leave out the unique row identifiers!
    .data$var_name_orig == "rowid" ~ "rowid", 
    TRUE ~ .data$var_label_orig 
  )) %>%
  mutate ( na_label_target = case_when(
    .data$na_label_orig %in% c("DK", "Refusal") ~ "Declined",
    TRUE ~ .data$na_label_target
  )) %>%
  mutate ( val_numeric_target = case_when(
    .data$val_label_orig == "No full-time education" ~ 14,
    .data$val_label_orig == "Still studying"         ~ 0,
    TRUE ~ .data$val_numeric_target
  )) %>%
  mutate ( na_numeric_target = case_when (
    .data$na_label_target == "Declined" ~ 99999,
    TRUE ~ NA_real_
  )) %>%
  mutate ( class_target = case_when (
    .data$var_name_target == "rowid" ~ 'character', 
    TRUE ~ "numeric"
  ))
```


```{r demography_crosstable}
set.seed(123456)
demography_crosstable %>% sample_n(5)
```
```{r demographyvars}
demography_vars <- crosswalk(survey_list     = cap_surveys, 
                             crosswalk_table =  demography_crosstable, 
                             na_values = NULL) %>%
  mutate ( is_student    = ifelse ( .data$age_education == 0, 1, 0), 
           age_education = ifelse (.data$age_education ==0, 
                                   .data$age_exact, 
                                   .data$age_education))
```

Now we have three variables including the `is_student` dummy:  

```{r summarydemography}
summary(demography_vars)
```

You can test yourself if people with `age_educaton`=20 are a homogeneous group, or they should be treated as a still studying group who have already been educated up to 20 years, and a group who had already left education at the age of 20. If the two groups are homogeneous, then you will no longer need to use the `is_student` variable.

```{r printdemography}
set.seed(1235)
demography_vars %>% sample_n(6) 
```

### Cultural Access & Participation Variables

The survey items that we are interested in are in the questionnaire block QB1 (in the 2013 questionnaire.)

- `How many times in the last twelve months have you…? [Been to a concert]`---basic (master English, identical to the Irish version.)
- `Au cours des 12 derniers mois, combien de fois … ? [Êtes-vous allé(e) à un concert]`---basic (master) French version
- `Hoeveel keer gedurende de afgelopen 12 maanden heeft u…? [een concert bijgewoond]`---Dutch localized to the Netherlands (may differ from Dutch localized to Belgium)
- `Az elmúlt tizenkét hónapban hány alkalommal ….? [Volt koncerten]`---localized Hungarian verison.


Our variables of interest are visiting frequencies to concerts and public libraries.  In the Eurobarometer CAP surveys, the answers have 

1.	`Never in the last twelve months` or `None` will be coded as 0 visits, and their factor label will be harmonized to `never`. In this case, the questionnaire reveals that whilst this was an *ex ante* harmonized question, the answers were not consistently labelled in the three files. We can safely bring these responses to the same value
2.	`1-2 times` will be coded to the programatically easier to use `1_2_times`, and it will get a numeric value of `1.5`.  Later, we will try to establish a better numeric value, now it is important that all responses labelled as `1-2 times` should have the same numeric code.
3.	`3-5 times` will be coded to the numeric value of `3.5` and `3_5_times`.
4.	`More than 5 times` will be `more_than_5` and have the numeric value of `6`.
5.	The special case of `DK` will be coded as `declined` (to anwer) and the *Inap...* labels to `inap`.  These are special codes in Eurobarometer meaning that the item is inappropriate, i.e. due to some filtering, this particular question was not asked from the respondent. While respondents who consciously decline an answer usually form a rather homogenous category of their own, the inappropriate answers are truly missing answers.  Their numeric representation will be both `NA` (f.e. they should be omitted from a numeric average.)

For code readability, we show the following data pipeline in sections, but you could of course create this code in a single expression---or create the resulting table in a spreadsheet application, if you are not aiming for full reproducibility in your research.

1. Let's create a search term for the metadata inventory:

```{r capvarscrosstable1}
search_term_labels <- c("rowid|cultural_activities_public_library|artistic_activities_sung|cultural_activities_concert|cultural_activities_freq_concert|artistic_activities_played_music_instr|artistic_activities_played_music_instrument
")
```

2. Create an empty schema crosswalk table:

```{r capvarscrosstable2}
cap_vars_crosstable <- cap_metadata %>%
  filter ( .data$var_name_orig == "rowid" | 
             grepl(search_term_labels, .data$var_label_orig) ) %>%
  crosswalk_table_create ()

set.seed(6)
cap_vars_crosstable %>%
  sample_n(6)
```

_The following steps can be made in a spreadsheet application, too, here we fill out the table programatically_:

3. Set the target variable names for the harmonization of the variables (columns) of the final, tidy, unified dataset.

```{r capvarscrosstable3}
cap_vars_crosstable <- cap_vars_crosstable %>%
  mutate ( var_name_target = case_when(
    .data$var_name_orig == "rowid" ~ "rowid", 
    grepl("concert", .data$var_label_orig)       ~ "visit_concert", 
    grepl("sung", .data$var_label_orig)          ~ "activity_sung", 
    grepl("played_music", .data$var_label_orig)  ~ "activity_played_music", 
    TRUE ~ .data$var_label_orig 
  ))
```

4. Define the special (missing value) labels. It is important to convert any data with such labels as the R special `NA_real_` value when you give a numeric representation to your data:

```{r capvarscrosstable4}
cap_vars_crosstable <- cap_vars_crosstable %>%
  mutate ( na_label_target = case_when(
    .data$na_label_orig %in% c("DK", "Refusal") ~ "declined",
    grepl("^Inap", .data$na_label_orig)         ~ "inap",
    .data$val_label_orig == "DK"                ~ "declined",
    TRUE ~ .data$na_label_target
  )) %>% 
  mutate ( na_numeric_target = case_when (
    .data$na_label_target == "declined" ~ 99999,
    .data$na_label_target == "inap"     ~ 99998,
    TRUE ~ NA_real_
  )) 
```

5. Define the labels in the normal value range. These will be the factor levels in a categorical representation of your data:

```{r capvarscrosstable5}
cap_vars_crosstable <- cap_vars_crosstable %>%
  mutate ( val_label_target = case_when (
    tolower(.data$val_label_orig) == "mentioned"             ~ "mentioned",
    tolower(.data$val_label_orig) == "not mentioned"         ~ "not_mentioned",
    tolower(.data$val_label_orig) == "1-2 times"             ~ "1_2_times", 
    tolower(.data$val_label_orig) == "3-5 times"             ~ "3_5_times",
    grepl("^more", tolower(.data$val_label_orig))            ~ "more_than_5", 
    grepl("never|^not\\s\\in|none", tolower(.data$val_label_orig))  ~ "never",
    !is.na(.data$na_label_target) ~  .data$na_label_target,
    # do not forget the rowid, it should not be labelled
    .data$var_name_target == "rowid" ~ NA_character_,
    TRUE ~ "coding_error"   # If this appears in your scheme crosswalk table, you have a problem.
  )) 
```

6. Define the numeric representation of your labels. These will be the factor levels in a numeric representation of your data:

```{r capvarscrosstable6}
cap_vars_crosstable <- cap_vars_crosstable %>%
  mutate ( val_numeric_target = case_when (
    .data$val_label_target == "never"         ~ 0,
    .data$val_label_target == "1_2_times"     ~ 1.5,
    .data$val_label_target == "3_5_times"     ~ 3.5,
    .data$val_label_target == "more_than_5"   ~ 6,
    .data$val_label_target == "mentioned"     ~ 1,
    .data$val_label_target == "not_mentioned" ~ 0,
    TRUE ~ .data$na_numeric_target
  ))

```

At last, we create two versions of the crosswalk table. 

One will rename the `visit_concert` variable to `fct_visit`, and give them the categorical representation. The `crosswalk()` function will use the retroharmonize `as_factor` method instead of base R's `as.factor` to properly treat missing values.

```{r capfct}
cap_vars_crosstable_fct <- cap_vars_crosstable %>%
  mutate ( var_name_target = ifelse(
           test = grepl("^visit|^artistic", .data$var_name_target), 
           yes  = paste0("fct_", .data$var_name_target), 
           no   = .data$var_name_target), 
           class_target = ifelse(
             test  = .data$var_name_target == 'rowid', 
             yes = "character", # the rowid remains a character
             no  = "factor" )
           )

cap_vars_fct <- crosswalk(
  crosswalk_table = cap_vars_crosstable_fct,  
  survey_list = cap_surveys)

set.seed(2022)
cap_vars_fct %>% sample_n(6)
```

The other will give these variable a numeric representation. The `crosswalk()` function will use the retroharmonize `as_numeric` method instead of base R's `as.numeric` to avoid unexpected coercion in the case of labelled missing cases.

```{r capnum}
cap_vars_crosstable_num <- cap_vars_crosstable %>%
   mutate( class_target = ifelse(
      test  = .data$var_name_target == 'rowid', 
      yes = "character",  # the rowid remains a character
      no = "numeric" )
  )

cap_vars_num <- crosswalk(
  crosswalk_table = cap_vars_crosstable_num,  
  survey_list = cap_surveys)

set.seed(2013)
cap_vars_crosstable_num %>% sample_n(6)
```

## Join the harmonized CAP dataset

Using [dplyr::left_join](https://dplyr.tidyverse.org/reference/mutate-joins.html) we join together by the survey `id` and the observation `rowid` the categorical and numeric representations of the CAP surveys, then add the demography variables, then add the weights.

```{r joinharmoinzed}
harmonized_cap_file <- cap_vars_fct %>%
  left_join ( cap_vars_num,    by = c("id", "rowid") ) %>%
  left_join ( demography_vars, by = c("id", "rowid") ) %>%
  left_join ( weight_vars,     by = c("id", "rowid") )
```


## Unit testing and corrections

The following simple logical tests should give an indication if our results are as expected.

### Weights

The average post-stratification weight must be 1 for `w` and `w1`.  The `wex` variable has no naturally defined mean value. 

```{r weighttest, eval=FALSE}
weight_vars %>%
  group_by ( .data$id ) %>%
  mutate ( across(starts_with("w"), function(x) ifelse(x==0, NA, x))) %>% 
  dplyr::summarise( across(starts_with('w'), mean, na.rm=TRUE))
```

The post-stratification weights have a very small deviation from zero.  Use these values when you calculate weighted averages.  The projection weight should be used when you calculate sums.

### Missing labels

All concert visits labelled as `declined` and `inap` must be coded to a numeric `NA` value.  Furthermore, there must not be other `NA` values than cases in the `ZA6925_v1-0-0` survey, where this question was not asked. For consistency, these observations should be also coded as `inap` in the factor representation. We use [dplyr](https://dplyr.tidyverse.org/index.html) for the final data wrangling, but of course, you can do this in [DT](https://cran.r-project.org/web/packages/DT/index.html) or base R if you prefer.

```{r correctmissings}
harmonized_cap_file %>% 
  select ( starts_with(c("id", "fct", "visit")) ) %>%
  filter ( is.na(.data$visit_concert) ) %>%
  distinct_all()
```

### Create a visiting binary variable

We create `is_visit_concert`, which takes the value of 1 if the person visited at least one concert in the previous 12 months, and 0 otherwise, retaining the missing values from declined answers.

```{r isvisit}
harmonized_cap_file <- harmonized_cap_file %>%
  mutate ( is_visit_concert = ifelse(.data$visit_concert > 1, 1, .data$visit_concert)) 
```

And checking if it is indeed smaller than the original (estimated) visiting frequency value.

```{r checkisvisit}
harmonized_cap_file %>% 
  group_by ( .data$id) %>%
  dplyr::summarise ( across(starts_with("visit"),     mean, na.rm=TRUE), 
                     across(starts_with("is_visit"),  mean, na.rm=TRUE), 
                     .groups = "keep")
```

## Exporting

You can find this harmonized dataset on Zenodo in the [Digital Music Observatory](https://zenodo.org/communities/music_observatory/) and the [Cultural Creative Sectors Industries Data Observatory](https://zenodo.org/communities/ccsi/) repositories. It was exported with the following code:


```{r savefile, eval=FALSE}
saveRDS(harmonized_cap_file,  file.path(tempdir(), "harmonized_cap_dataset.rds"))
write.csv(harmonized_cap_file,  
          file.path(tempdir(), "harmonized_cap_file_20220129_doi_10_5281_zenodo_5917742.csv"), 
          row.names = FALSE)
```

Daniel Antal. (2022). Harmonized Cultural Access & Participation Dataset for Music (Version 20220002) [Data set]. Zenodo. [https://doi.org/10.5281/zenodo.5917742](10.5281/zenodo.6606867)

Daniel Antal. (2022). Creating a Harmonized Cultural Access & Participation Dataset for Music (Version 20220129) [Data set]. Zenodo. [10.5281/zenodo.5917714](https://doi.org/10.5281/zenodo.5917714)

