---
title: "Survey Harmonization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Survey Harmonization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(retroharmonize)
```

Survey data harmonization refers to procedures that improve the data comparability or the inferential capacity of multiple surveys. *Ex ante* survey harmonization refers to planning and design steps to make sure that not yet answerred questionnaires can be better compared, or data derived from them joined, integrated. Such procedures include the harmonization of the questionnaire, the harmonization of the sample design, and other aspects of carrying out multiple surveys. *Ex post* or *retrospective harmonization* refers to procedures to data that has been derived from surveys---i.e., survey that have been carried out. 

Naturally, better *ex ante* harmonization makes eventual data integration or data comparison easier; yet often we can still harmonize retrospectively survey data that has not been carefully pre-harmonized before respondents have answered the questionnaire items.

Our aim with the `retroharmonize` R package is to provide assistance to a reproducible research workflow in carrying out important computational aspects of retrospective survey harmonization. 

Let's start with a very simple example.

```{r simple-example-def}
library(labelled)
survey_1 <- data.frame(
  sex = labelled(c(1,1,0, NA_real_), c(Male = 1, Female = 0))
)
attr(survey_1, "id") <- "Survey 1"

survey_2 <- data.frame(
  gender = labelled(c(1,3,9,1,2), c(male = 1, female = 2, other = 3, declined = 9))
)
attr(survey_2, "id") <- "Survey 2"
```


```{r ex1}
library(dplyr, quietly = TRUE)
survey_1 %>%
  mutate ( sex_numeric = as_numeric(.data$sex), 
           sex_factor = as_factor(.data$sex))
```

## Tasks in the harmonization workflow

The ordering of the survey harmonization workflow is flexible, and it is likely that even the same researcher would choose a different workflow in the case of smaller, simpler harmonization tasks and more complex harmonization tasks.

The data science aspect of a successful survey harmonization task is the creation of a consistent data frame that contains harmonized information from multiple surveys. It practically means that questionnaire items are mapped into variables with a consistent numerical coding, descriptive metadata (variable and value labels) and a consistent handling of missing and special values. In practice, this is a very laborous tasks, when surveys are conducted in different years, saved in different file formats with a different metadata structure, missing and special values are handled differently, and the metadata contains potentially different natural language descriptions or spelling.

`Survey 1` labels the sex of respondents as `Male` and `Female`, and has cases that are neither `Male` or `Female`, but we do not know why.

```{r example2}
survey_2 %>%
  mutate ( gender_numeric = as_numeric(.data$gender), 
           gender_factor  = as_factor(.data$gender))
```

`Survey 2` records gender, which contains the same information as sex in `Survey 1` (`Male` and `Female`), but allows people to identify as `Other`, and labels cases when people decline to identify with any of these three categories.

In practice, you want to end up with the following joined representation of your survey:

```{r manually-joined}
survey_joined  <- data.frame(
  id = c(1,2,3,4,1,2,3,4,5),
  survey = c(rep(1,4), rep(2, 5)),
  gender = labelled(c(1,1,0,9, 1,3,9,1,0), c(male = 1, female = 0, other = 3, declined = 9))
)

survey_joined %>%
  mutate ( id = paste0("survey_", .data$survey, "_", .data$id),
           gender_numeric = c(1,1,0,NA_real_, 1,3,NA_real_,1,0), 
           gender_factor  = as_factor(.data$gender), 
           is_female = ifelse (.data$gender_numeric == 0, 1, 0))
```


1.  Harmonization of concepts
    *  Create a mental map of the measured concepts that needs to be harmonized. Which variables contain sufficiently similar information that can be harmonized? In our simple example, we want to harmonize a binary sex with missing cases and a four-level categorical variable on gender identification, and concatenate the harmonized vectors by binding or joining the Survey 1 and Survey 2 data frames.
    *  Our metadata function help mapping the information stored in the file representations of multiple surveys. We want to create a simple inventory of numerical codes, value ranges, missing cases and variable labels.
2.  Variable names
    *  Data measuring sufficiently similar concepts, i.e. data that can be harmonized, is stored in variables that have the same name in different data frames representing the survey, therefore they can be bind or joined together. We want to join or bind by rows `survey_1` with `survey_2`, or, we want to concatenate `survey_1$sex` with `survey_2$gender`.
    *  Descriptive metadata about the variable, such as "variable labels" in SPSS files, is recorded for documentation, and if needed, harmonized across surveys. In SPSSS,  `survey_1$sex` may come with a variable label something like *SEX OF RESPONDENT*, and `survey_2$gender` may be labelled as *GENDER IDENTIFICATION*. This label should be harmonized to *Sex or gender or the respondent*. 
3.  Variable coding and labels
    * Variables recording or measuring the same concept, such as the gender of the respondent, are coded exactly the same way, for example, females with 0, males with 1, non-binary respondents with 3, and people declining to reveal their gender with 9. This means that observations in `survey_2$gender` coded with a numeric 2 must be changed to a numeric 0.
    * Variable labels are used consistently and in exactly the same way, i.e. `survey_1$sex` *Female* respondents and `survey_2$gender` *female* respondents will be consistently labelled as *female*.
4.  Variable types
    * Variables recording or measuring the same concept are stored in exactly the same R type, and they can be consistently concatenated across surveys, or they can be subsetted, cross-cutting surveys, for example, all female respondents from Survey 1 and Survey 2 can be subsetted into a female vector.
    * The labelled class of [labelled](https://cran.r-project.org/web/packages/labelled/vignettes/intro_labelled.html) is not sufficiently strict, because it allows inconsistent special (missing) values.  Our interited [labelled_spss_survey](https://retroharmonize.dataobservatory.eu/reference/labelled_spss_survey.html) consistently contains codes, labels, missing ranges and missing values, and therefore it can be concatenated.
    * The numeric or factor representation of `survey_1$sex` and `survey_2$gender` can be technically concatenated, but before harmonization this will create logical errors, because females will be either coded with 0 or with 2. The `as_numeric()` and `as_factor()` methods of our [labelled_spss_survey](https://retroharmonize.dataobservatory.eu/reference/labelled_spss_survey.html)  class handle consistency issues.
5.  Reproducibility
    * The revision, checking, external review and audit of the data requires that the steps can be replicated by a third party. This requires a documentation of the harmonization steps, i.e., 1=Women in Survey 1, and 0=female in Survey 2 became 0=females in the harmonized dataset.
    * Our [survey](https://retroharmonize.dataobservatory.eu/reference/survey.html) class is derived from [tibble](https://tibble.tidyverse.org/), the modernized version of the base `data.frame()`. It contains various descriptive metadata about the survey among attributes.
    
The joining of the not harmonized datasets results in the following data frame.
    
```{r}
library(dplyr)

survey_1 %>%
  mutate ( survey = 1, 
           sex_numeric = as_numeric(.data$sex), 
           sex_factor = as_factor(.data$sex)) %>%
  full_join( 
    survey_2 %>%
      mutate ( survey = 2, 
               gender_numeric = as_numeric(.data$gender), 
               gender_factor  = as_factor(.data$gender))
    )

```  
    
Performing only variable harmonization yields to a data frame that has the correct dimensions, but it is not usable for statistical analysis. 

    
```{r naive-join}
library(dplyr)

survey_var_harmonized <- survey_1 %>%
  rename ( gender = .data$sex ) %>%
  mutate ( survey = 1, 
           gender_numeric = as_numeric(.data$gender), 
           gender_factor  = as_factor(.data$gender)) %>%
  full_join( 
    survey_2 %>%
      mutate ( survey = 2, 
               gender_numeric = as_numeric(.data$gender), 
               gender_factor  = as_factor(.data$gender)), 
     by = c("gender", "survey", "gender_numeric", "gender_factor")
    )

```      

Apart from the simple, descriptive variable of the survey identification, non of the descriptive statistics are meaningful.    

```{r}
summary(survey_var_harmonized)
```
The variable labels must be harmonized for a successful factor representation.  The numerical coding must be harmonized, and the missing cases must be consistently handled to acheive any useful numerical representation. 

```{r}
survey_joined %>%
  mutate ( id = paste0("survey_", .data$survey, "_", .data$id),
           gender_numeric = c(1,1,0,NA_real_, 1,3,NA_real_,1,0), 
           gender_factor  = as_factor(.data$gender), 
           female_ratio = ifelse (.data$gender_numeric == 0, 1, 0)) %>%
  summary()
```
## How we help harmonization

1. The *data importing* functions make sure that survey data and metadata are carefully translated to R data classes and variable types.  

2. The *metadata functions* help the analysis, normalization and joining of the metadata aspects (variable and value labels, original variable names, unique identifiers) across surveys.

3. *Harmonization functions* help the harmonization of responses to questionnaire items, i.e. making sure that coded values, the labelling of values, and missing data are handled consistently across multiple surveys.

Our package was tested on multiple, international, harmonized surveys, particularly the Eurobarometer, the Afrobarometer and the Arab Barometer survey programs. Different users, and different task call for different workflows.  We created a number of helper functions to assist various workflows.  
